### 2020-09-13 paper/coloringAppendix/algVII.tex
Plots and appendix

\begin{algorithm}
  \SetKwFunction{FindSpecialK}{\textsc{Find-Special-K4-Strip-System}}
	\Indm\nonl\FindSpecialK{$G, J, (S, N), m$}\\
  \KwData{$G$ -- square-free, Berge graph \newline
    $J$ -- a $3$-connected graph with appearance in $G$ \newline
    $(S, N)$ -- a $J$-strip system \newline
    $m \in G \setminus (S, N)$ that is major w.r.t. some choice of rungs of $(S, N)$ but not w.r.t. $(S, N)$
  }
  \KwResult{$(S', N')$ -- A special $K_4$-strip system, or \newline
  $(S'', N'')$ -- a bigger $J$-strip system
  }
  \Indp	
  $X \gets N(m)$\;
  $M \gets$ vertices of $G\setminus V(S,N)$ major w.r.t. $(S,N)$\;
  $M^* \gets$ vertices of $G\setminus V(S,N)$ major w.r.t. some choice of rungs\;
  \Assert[SPGT 8.4]{$J = K_4$}
  $V(J) \gets [4]$\;
  $\forall_{i\neq j \in [4]}$: choose rungs $R_{ij}, R_{ij}$, forming line graphs $L(H)$ and $L(H')$ so that $X$ 
  saturates $L(H)$ but does not saturate $L(H')$\;
  \Assert{$R_{ij} \neq R'_{ij} \if and only if \{i, j\} = [2]$}
  $r_{ij}, r_{ji} \gets$ ends of each $R_{ij}$\;
  $r'_{ij}, r'_{ji} \gets$ ends of each $R'_{ij}$\;
  $\forall_{i \in [4]} T_i \gets \{r_{ij}, j \in [4] \setminus \{i\} \}$\;
  $\forall_{i \in [4]} T'_i \gets \{r'_{ij}, j \in [4] \setminus \{i\} \}$\;
  \Assert{$X$ has at least two members in each $T_1, \ldots T_4$}
  \Assert{There is $T'_i$ that contains at most one member of $X$}
  \Assert{$T_3 = T'_3, T_4 = T'_4$}
  relabel 1 and 2 in $J$, so that $|X \cap T_1| = 2$ and $|X \cap T'_1| = 1$\;
  \Assert{$r_{12} \in X, r'_{12} \notin X$}
  \Assert{$r_{13} \in X \xor r_{14} \in X$}
  relabel 3 and 4 in $J$, so that $r_{13} \in X, r_{14} \notin X$\;
  \Assert[6.3.(3)]{$R_{34}$ is even and $[X \cap V(L(H'))] \setminus V(R_{34}) = \{r_{31}, r_{32}, r_{41}, r_{42} \}$}
  \Assert[6.3.(4)]{$R_{14}$ has odd length, $r_{21} \in X$}
  \Assert[6.3.(5)]{$R_{12}$ has length $0$, every $12$-rung has even length}
  \Assert[6.3.(6)]{$R_{24}$ has length $0$ and $R_{23}$ has odd length}
  \Assert[6.3.(7)]{Every $34$-rung has non-zero even length}
  $\forall_{i \neq j \in [4]} O_{ij} \gets$ set of vertices that are not major w.r.t $L(H')$ and
  are complete to $(T'_i \cup T'_j) \setminus R'_{ij}$\;
  \Assert{$r_{12} = r_{21} \in O_{12}$, $m \in O_{34}$}
  \Assert{$O_{34} = M^* \setminus M$}
  $(S', N') \gets$ strip system obtained from $(S, N)$ by replacing $S_{12}$ with $S_{12} \setminus O_{12}$\;
  \If{$\exists_{rung R}: $ adding $R$ to $S'_{12}$ produces enlargement of $(S, N)$}{
    \Return $(S'', N'')$ -- an enlargement of $(S, N)$\;
  }
  \Else{
    \Return $(S', N')$ -- a special $K_4$ strip system\;
  }
  \vspace{-10cm}
\end{algorithm}

>>>>

\begin{mynameforalgorithm}
  \SetKwFunction{FindSpecialK}{\textsc{Find-Special-K4-Strip-System}}
	\Indm\nonl\FindSpecialK{$G, J, (S, N), m$}\\
  \KwData{$G$ -- square-free, Berge graph \newline
    $J$ -- a $3$-connected graph with appearance in $G$ \newline
    $(S, N)$ -- a $J$-strip system \newline
    $m \in G \setminus (S, N)$ that is major w.r.t. some choice of rungs of $(S, N)$ but not w.r.t. $(S, N)$
  }
  \KwResult{$(S', N')$ -- A special $K_4$-strip system, or \newline
  $(S'', N'')$ -- a bigger $J$-strip system
  }
  \Indp	
  $X \gets N(m)$\;
  $M \gets$ vertices of $G\setminus V(S,N)$ major w.r.t. $(S,N)$\;
  $M^* \gets$ vertices of $G\setminus V(S,N)$ major w.r.t. some choice of rungs\;
  \Assert[SPGT 8.4]{$J = K_4$}
  $V(J) \gets [4]$\;
  $\forall_{i\neq j \in [4]}$: choose rungs $R_{ij}, R_{ij}$, forming line graphs $L(H)$ and $L(H')$ so that $X$ 
  saturates $L(H)$ but does not saturate $L(H')$\;
  \Assert{$R_{ij} \neq R'_{ij}$ if and only if $\{i, j\} = [2]$}
  $r_{ij}, r_{ji} \gets$ ends of each $R_{ij}$\;
  $r'_{ij}, r'_{ji} \gets$ ends of each $R'_{ij}$\;
  $\forall_{i \in [4]} T_i \gets \{r_{ij}, j \in [4] \setminus \{i\} \}$\;
  $\forall_{i \in [4]} T'_i \gets \{r'_{ij}, j \in [4] \setminus \{i\} \}$\;
  \Assert{$X$ has at least two members in each $T_1, \ldots T_4$}
  \Assert{There is $T'_i$ that contains at most one member of $X$}
  \Assert{$T_3 = T'_3, T_4 = T'_4$}
  relabel 1 and 2 in $J$, so that $|X \cap T_1| = 2$ and $|X \cap T'_1| = 1$\;
  \Assert{$r_{12} \in X, r'_{12} \notin X$}
  \Assert{$r_{13} \in X \xor r_{14} \in X$}
  relabel 3 and 4 in $J$, so that $r_{13} \in X, r_{14} \notin X$\;
  \Assert[6.3.(3)]{$R_{34}$ is even and $[X \cap V(L(H'))] \setminus V(R_{34}) = \{r_{31}, r_{32}, r_{41}, r_{42} \}$}
  \Assert[6.3.(4)]{$R_{14}$ has odd length, $r_{21} \in X$}
  \Assert[6.3.(5)]{$R_{12}$ has length $0$, every $12$-rung has even length}
  \Assert[6.3.(6)]{$R_{24}$ has length $0$ and $R_{23}$ has odd length}
  \Assert[6.3.(7)]{Every $34$-rung has non-zero even length}
  $\forall_{i \neq j \in [4]} O_{ij} \gets$ set of vertices that are not major w.r.t $L(H')$ and
  are complete to $(T'_i \cup T'_j) \setminus R'_{ij}$\;
  \Assert{$r_{12} = r_{21} \in O_{12}$, $m \in O_{34}$}
  \Assert{$O_{34} = M^* \setminus M$}
  $(S', N') \gets$ strip system obtained from $(S, N)$ by replacing $S_{12}$ with $S_{12} \setminus O_{12}$\;
  \If{$\exists_{rung R}: $ adding $R$ to $S'_{12}$ produces enlargement of $(S, N)$}{
    \Return $(S'', N'')$ -- an enlargement of $(S, N)$\;
  }
  \Else{
    \Return $(S', N')$ -- a special $K_4$ strip system\;
  }
  \vspace{-10cm}
\end{mynameforalgorithm}

### 2020-09-13 paper/experiments.tex
Plots and appendix

We ran tests on X classes of perfect graphs, see \Cref{fig:graphClasses} for some examples:
\begin{itemize}
  \item Random perfect graphs,
  \item Random bipartite graphs,
  \item Line graphs of random bipartite graphs,
  \item Lattice graphs (\cref{fig:lattice}),
  \item Rook graphs (\cref{fig:rook}),
  \item Knight graphs (\cref{fig:knight}),
  \item Hypercube graphs (\cref{fig:hypercube}),
  \item Split graphs (\cref{fig:split}).
\end{itemize}

In later paragraphs we describe each class one by one and benchmark our implementations on them.

@@

\paragraph{Random perfect graphs}
Our first method of generating graphs is to set $|V|$ and for each pair of vertices $u, v$ let $uv$ be the edge with probability of $1/2$. Then we check if generated graph is perfect and continue as long as we don't get sufficient number of perfect graphs.

First, we note that this method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favour the na\"ive algortihm, because there is very low probability of long chordless paths to appear (each chord has a $1/2$ chance to appear).

@@

\TODO{Better plot by a lot WIP}

\begin{figure}
  \centering
  \input{plots/perf.t.in.out.csv.lines.pgf}
  \label{plot:perfLines}
  \caption{Random perfect graphs}
\end{figure}

\paragraph{Random bipartite graphs}
\TODO{TODO}

\paragraph{Line graphs of random bipartite graphs}
\TODO{data with bigger N}
Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. We repeated the process until we had sufficieng number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see GPU gives us much better improvement than in random graphs. This is due to the fact, that testing near cleaners takes over 80\% of the time for bigger graphs (\Cref{plot:perf2Det}).

\begin{figure}
  \centering
  \input{plots/perf2.t.in.out.csv.lines.pgf}
  \label{plot:perf2Lines}
  \caption{Line graphs of random bipartite graphs}
\end{figure}

\paragraph{Lattice graphs}

Next, we turn our attention to graphs generated on a grid, or a checker board (see later paragraphs). First, we take lattice graphs. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \cref{fig:lattice} for an example of a $4 \times 4$ lattice graph.

@@

With larger $|V|$ the speedup gained from utilizing GPU is also more significant. For $|V| = 78$ the running time for the whole algorithm is almost 3x better.

\begin{figure}
  \centering
  \input{plots/grid6by5to11.t.in.out.csv.lines.pgf}
  \label{plot:gridLines}
  \caption{Lattice graphs}
\end{figure}

\paragraph{Rook graphs}

If we imagine our grid to be a chessboard we can define a Rook graph to have an edge, when the rook from chess can move from one vertex to another in one move. See \cref{fig:rook} for an example of $4 \times 4$ Rook graph.

@@

\begin{figure}
  \centering
  \input{plots/rookGraph5by4to7.t.in.out.csv.lines.pgf}
  \label{plot:rookLines}
  \caption{Rook graphs}
\end{figure}

\paragraph{Knight graph}

A knight graph is constructed in a similar way to rook graph, but by taking potential moves of a knight.

\TODO{add 800s result for naive}

\begin{figure}
  \centering
  \input{plots/knightGraph8by4to8.t.in.out.csv.lines.pgf}
  \label{plot:knightLines}
  \caption{Knight graphs}
\end{figure}

\paragraph{Hypercube graphs}

@@

Hypercube graphs tell a similar story to lattice graphs and knight graphs. Na\"ive is very fast at first, but becomes unbearably slow for $|V|$ higher than 55. Unfortunately, CCLSV doesn't fare much better, although it grows slower. Again, the speedup of GPU CCLSV is noticeable, being around 4.8x for $|V| = 55$.

\begin{figure}
  \centering
  \input{plots/hypercubes20to55.t.in.out.csv.lines.pgf}
  \label{plot:hypercubeLines}
  \caption{Hypercube graphs}
\end{figure}

\paragraph{Split graphs}

@@

There are no long chordless paths here, so na\"ive fares great (\cref{plot:splitLines}). What is interesting, when looking at detailed times of GPU CCLSV, we can see that checking the existence of simple forbidden structures takes more than 30\% of the time. This opens a possibility of further GPU speedups, not explored by us.

\begin{figure}
  \centering
  \input{plots/split20to50.t.in.out.csv.lines.pgf}
  \label{plot:splitLines}
  \caption{Split graphs}
\end{figure}

\begin{figure}
  \centering
  \input{plots/detailed.csv.detailed.pgf}
  \label{plot:splitLines}
  \caption{\the\textwidth \the\textheight}
\end{figure}

@@

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.perf.t.in.out.csv.lines.pgf}
    \caption{Overall running times}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.perf.t.in.out.csv.detailed.pgf}
    \caption{Percentage of time taken be each part}
  \end{subfigure}
  \caption{Random perfect graphs}
\end{figure}

\begin{figure}
  \TODO{TODO}
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf2.t.in.out.csv.lines.pgf}
  %   \caption{Overall running times}
  %   \label{plot:colorPerfLines}
  % \end{subfigure}%
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf2.t.in.out.csv.detailed.pgf}
  %   \caption{Percentage of time taken be each part}
  %   \label{plot:colorPerfDet}
  % \end{subfigure}
  \caption{Line graphs of random bipartite graphs}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.grid5by4to9.t.in.out.csv.lines.pgf}
    \caption{Overall running times}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.grid5by4to9.t.in.out.csv.detailed.pgf}
    \caption{Percentage of time taken be each part}
  \end{subfigure}
  \caption{Lattice graphs}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.rookGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Overall running times}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.rookGraph6by4to8.t.in.out.csv.detailed.pgf}
    \caption{Percentage of time taken be each part}
    \label{plot:colorRookDet}
  \end{subfigure}
  \caption{Rook graphs}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.knightGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Overall running times}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.knightGraph6by4to8.t.in.out.csv.detailed.pgf}
    \caption{Percentage of time taken be each part}
  \end{subfigure}
  \caption{Knight graphs}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.hypercube20to40.t.in.out.csv.lines.pgf}
    \caption{Overall running times}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.hypercube20to40.t.in.out.csv.detailed.pgf}
    \caption{Percentage of time taken be each part}
  \end{subfigure}
  \caption{Hypercube graphs}
\end{figure}

\begin{figure}
  \TODO{TODOa}
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf2.t.in.out.csv.lines.pgf}
  %   \caption{Overall running times}
  %   \label{plot:colorPerfLines}
  % \end{subfigure}%
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf2.t.in.out.csv.detailed.pgf}
  %   \caption{Percentage of time taken be each part}
  %   \label{plot:colorPerfDet}
  % \end{subfigure}
  \caption{Split graphs}
\end{figure}

>>>>

Below, we first present our data sets then overall time results on them and lastly a plot breaking down what parts of CCLSV and GPU CCLSV constitute their overall running time.

We ran tests on nine classes of perfect graphs.
\begin{itemize}
  \item perfect Erd\H{o}s--Rényi graphs,
  \item Random bipartite graphs,
  \item Line graphs of random bipartite graphs,
  \item Lattice graphs (\cref{fig:lattice}),
  \item Rook graphs (\cref{fig:rook}),
  \item Knight graphs (\cref{fig:knight}),
  \item Hypercube graphs (\cref{fig:hypercube}),
  \item Split graphs (\cref{fig:split}),
  \item Full binary trees.
\end{itemize}

@@

In \Cref{fig:graphClasses} we present small examples of some of the above classes. Let us describe them briefly.

An \emph{Erd\H{o}s--Rényi} model is a method of generating random graphs. Every possible edge is included in a graph with probability $1/2$. We generated them as long as we found enough perfect graphs for each $|V|$ we wanted to test. \todo{implement generating perfect graphs by removing long cycles} This method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favour the na\"ive algortihm, because there is very low probability of long paths to appear.

Next, we generate bipartite graphs in similar manner -- each edge between two equal sized node sets has a probability of $1/2$ to appear. We also generate line graphs of bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

There are also a few interesting classes of perfect graphs that appear on a grid, or a checker board. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \Cref{fig:lattice} for an example of a $4 \times 4$ lattice graph. If we imagine our grid to be a chessboard we can define a rook graph to have an edge, when the rook from chess can move from one vertex to another in one move. See \Cref{fig:rook} for an example of $4 \times 4$ rook graph. In similar manner we can define a knight graph (\Cref{fig:knight}).

\TODO{split better granulity and detailed}
Split graphs we use are unions of cliques and independent sets of same size, with some edges between them. In generating them, each edge between a vertex from a clique and a vertex from an independent set has had $1/2$ chance to appear.

Finally, we benchmark on full binary trees to see how CCLSV works with a tree as an input.

\subsubsection{Time results}

% \begin{figure}
%   \centering
%   \input{plots/perf.t.in.out.csv.lines.pgf}
%   \caption{Random perfect graphs}
%   \label{plot:perfLines}
% \end{figure}

\begin{figure}
  \centering
  \input{plots/biparite.t.in.out.csv.lines.pgf}
  \caption{TODO: redo with more data Random bipartite graphs}
  \label{plot:perf2Lines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/perf2.t.in.out.csv.lines.pgf}
  \caption{Line graphs of random bipartite graphs}
  \label{plot:perf2Lines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/grid6by5to11.t.in.out.csv.lines.pgf}
  \caption{Lattice graphs}
  \label{plot:gridLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/rookGraph5by4to7.t.in.out.csv.lines.pgf}
  \caption{Rook graphs}
  \label{plot:rookLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/knightGraph8by4to8.t.in.out.csv.lines.pgf}
  \caption{Knight graphs}
  \label{plot:knightLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/hypercubes20to55.t.in.out.csv.lines.pgf}
  \caption{Hypercube graphs}
  \label{plot:hypercubeLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/split20to50.t.in.out.csv.lines.pgf}
  \caption{Split graphs}
  \label{plot:splitLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/fullBinary20to100.t.in.out.csv.lines.pgf}
  \caption{Full binary trees}
  \label{plot:splitLines}
\end{figure}

\begin{figure}
  \centering
  \input{plots/detailed.csv.detailed.pgf}
  \caption{Detailed running times}
  \label{plot:splitLines}
\end{figure}

@@

Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see na\"ive algorithm being much faster on the majority of the tests, although its faster growth allows GPU CCLSV to be faster on the biggest tests.

We turn our attention to graphs generated on a grid, or a checker board (see later paragraphs). First, we take lattice graphs. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \cref{fig:lattice} for an example of a $4 \times 4$ lattice graph.

@@

If we imagine our grid to be a chessboard we can define a rook graph to have an edge, when the rook from chess can move from one vertex to another in one move. See \cref{fig:rook} for an example of $4 \times 4$ Rook graph.

@@

@@

Hypercube graphs tell a similar story to lattice graphs and knight graphs. Na\"ive algorithm is very fast at first, but becomes unbearably slow for $|V|$ higher than 55. Unfortunately, CCLSV doesn't fare much better, although it grows slower. Again, the speedup of GPU CCLSV is noticeable, being around 4.8x for $|V| = 55$.

@@

There are no long paths here, so na\"ive algorithm works great (\cref{plot:splitLines}). What is interesting, when looking at detailed times of GPU CCLSV, we can see that checking the existence of simple forbidden structures takes more than 30\% of the time. This opens a possibility of further GPU speedups, not explored by us.

@@

\begin{figure}
  % \begin{subfigure}{.5\textwidth}
  %   \centering
  %   \input{plots/color.perf.t.in.out.csv.lines.pgf}
  %   \caption{Random perfect graphs}
  % \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Random perfect graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Line graphs of random perfect graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.grid5by4to9.t.in.out.csv.lines.pgf}
    \caption{Lattice graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.rookGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Rook graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{plots/color.knightGraph6by4to8.t.in.out.csv.lines.pgf}
    \caption{Knight graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering%
    \input{plots/color.hypercube20to40.t.in.out.csv.lines.pgf}
    \caption{Hypercube graphs}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Split graphs}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    TODO
    \caption{Full binary trees}
  \end{subfigure}
\end{figure}

\begin{figure}
  \centering
  \input{plots/color.detailed.csv.detailed.pgf}
  \caption{Detailed times of coloring}
\end{figure}
